\section{Data Structure}

\subsection{Array}
Array is a collection of items of the same variable type that are stored
at contiguous memory location. It is one of the most popular and simple
data structures used in programming.
\textbf{Basic terminologies of Array}
\begin{itemize}
\item Array Index: In an array, elements are identified by their
indexes. Array index starts from 0.
\item Array Element: Elements are items stored in an array and can be
accessed by their index.
\item Array Length: The length of an array is determined by the number
of elements it can contain.
\end{itemize}
\textbf{Memory representation of Array}
In an array, all the elements are stored in contiguous memory locations.
So, if we intitialize an array, of elements.
\textbf{Purpose of arrays}
Assume there is a class of five students and if we have to keep records
of their marks and examination then, we can do this by declaring five
variables individual and keeping track of records but what if the number
of students becomes very large, it would be challenging to manipulate
and maintain the data.

What it means is that, we can use normal variables (v1, v2, v3,...)
when we have a small number of objects. But if we want to store a large
number of instances, it becomes difficult to manage them with normal
variables. The idea of an array is to represent many instances in one
variable.

\textbf{Types of Arrays}
\begin{itemize}
    \item On the basis of Size
    \item On the basis of Dimensions
\end{itemize}

\textbf{Types of Arrays on the basis of Size:}
\begin{enumerate}
    \item Fixed Size Arrays:
        We Cannot alter or update the size of this array. Here only a
        fixed size (i,e. the size that is mentioned in square bracketes
        []) of memory will be allocated for storage. In case, we don't
        know the size of the array then if we declare a larger size and
        store a lesser number of elements will result in a wastage of
        memory or we declare a lesser size than the number of elements
        then we won't get enough memory to store all the elements. In
        such cases, static memory allocation is not preferred.
    \item Dynamic Sized Arrays:
        The size of the array changes as per user requirements during
        execution of cod so the coders do not have to warry about sizes.
        They can add and removed the elements as per the need. The
        memory is mostly dynamically allocated and de-allocated in these
        arrays.
\end{enumerate}

\textbf{Types of Arrays on the basis of Dimensions:}
\begin{enumerate}
    \item One-dimensional Array: You can imagine a 1d array as a row,
        where elements are stored one after another.
    \item Multi-dimensional Array: A multi -dimensional array is an
        array with more than one dimension. We can use multidimensional
        array to store complex data in the form of tables, etc. We can
        have 2D arrays, 3D arrays, 4D arrays and so on.
        \begin{enumerate}
            \item Two-Dimensional Array(Matrix): 2D Mulitdimensional
                arrays can be considered as an array of arrays or as a
                matrix consisting of rows and columns.
            \item Three-Dimensional Array: A 3D Mulitdimensional array
                contains three dimensions, so it can be considered an
                array of two-dimensional arrays.
        \end{enumerate}
\end{enumerate}
